---
title: "Homework 2"
author: 'BBIC: Bovee, Bichay, Iseneker, Coeman'
date: "2/2/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in the polity_vectors.rda file.

**Note: you will need to have your working directory set in order for this to load**

```{r}
  load("polity_vectors.rda")
```

```{r}
democ_factor_2016 <- democ_2016
democ_factor_2016[which(democ_2016 ==0)] <- 'Not a democracy'
democ_factor_2016[which(democ_2016 %in% 1:3)] <- 'Weak democracy'
democ_factor_2016[which(democ_2016 %in% 4:6)] <- 'Middle democracy'
democ_factor_2016[which(democ_2016 %in% 7:10)] <- 'High democracy'
```

Here, we classify democ_factor_2016 into four different types of democracy.

```{r}
democ_factor_2016 <- factor(democ_factor_2016,levels=c("Not a democracy", 
                                                        "Weak democracy", 
                                                        "Middle democracy", 
                                                        "High democracy"))
```

We use the factor function and assign four ordered levels to democ_factor_2016 here, going from least democratic (not a democracy) to most democratic (high democracy)

```{r}
levels(democ_factor_2016)[levels(democ_factor_2016) %in% c("Weak democracy", "Middle democracy")] <- "ambiguous"


attributes(democ_factor_2016)
```

We collapse 'weak democracy' and 'middle democracy' to be one level: ambiguous. 
