---
title: "Homework 3"
author: 'BBIC: Bovee, Bichay, Iseneker, Coeman'
date: "2/8/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1
### OLS function

First we simulate data for our LM 

```{r}
### simulate data

library(MASS)
set.seed(6886)
data = mvrnorm(
    n=100, mu=c(-2, 3), 
    Sigma=matrix(c(8,3,3,2),nrow=2,ncol=2) 
)
colnames(data) = c('y','x')
```

Here we create the function that calculates an OLS model manually. Comments within explain each step.  

```{r}
### create function

my_lm <- function(q, p){


# load data

x = as.matrix(cbind(1,p))
y = as.matrix(q)

# calculate b

beta = solve( t(x) %*% x) %*%t (x) %*% y

# calculate vcv and se

yhat <- x %*% beta   # predicted values

ssr <- sum((yhat - y)^2)   # sum of squared residuals

# set N = number of observations; k = number of variables (incl. intercept)

N <- nrow(y)
k <- ncol(x)

sigma2 <- (ssr/(N-k))   # variance

vcv <- (sigma2)*(solve( t(x) %*% x ))   # vcv

se <- sqrt(diag(vcv))

# output results

model_summary <- matrix(c(beta, se), nrow=2, ncol=2, 
                        dimnames=list( c("b0", "b1"),
                            c("Coef", "S.E."))
                        )

return(model_summary)

}

my_lm(data[,"y"], data[,"x"])

```



# Problem 2
### "Create this" matrix, from slide 37

Use the rep function to fill the data in the data frame. 

``` {r}
clever_df <- data.frame(Patient = c(1:16), 
                          Gender = rep(c("Male","Female"), each = 8),
                          Treatment1 = rep(c("Yes", "No"), each = 4),
                          Treatment2 = rep(c("Yes", "No"), each = 2),
                          Treatment3 = rep(c("Yes", "No"))
                          )
clever_df
```




# Problem 3
### Polity matrix and summary stats








# Problem 4
### Merging the other variables






# Problem 5
### Hadley problems

Data structures:
  Vectors:
1. What are the six types of atomic vector?   
logical, integer, double, character, complex and raw
How does a list differ from an atomic vector?
lists can include any element type, including other lists
  
2. What makes is.vector() and is.numeric() fundamentally different to is.list() and is.character()?
is.vector and is.numeric do not test for a specific type of element 

3. Test your knowledge of vector coercion rules by predicting the output of the following uses of c():
c(1, FALSE) <- double
c("a", 1) <- character
c(list(1), "a") <- list
c(TRUE, 1L) <- integer

4. Why do you need to use unlist() to convert a list to an atomic vector? Why doesn’t as.vector() work?
Lists are vectors, but not atomic vectors, so as.vector doesn't do anything as it is coercing a vector to a vector.  

5. Why is 1 == "1" true? Why is -1 < FALSE true? Why is "one" < 2 false?
1 == "1": the numeric is coerced into a character - moving up from a less flexible element to more
-1 < F: FALSE is coerced into an integer, become 0, making the statement -1 < 0
"one" < 2: the numeric is coerced into a character and "one" < "2" is then FALSE

6. Why is the default missing value, NA, a logical vector? What’s special about logical vectors? (Hint: think about c(FALSE, NA_character_).)
Logical is the least flexible type, so NA are logical so they don't lose any information when coerced. 

  Attributes:
1. An early draft used this code to illustrate structure():
structure(1:5, comment = "my attribute")
## [1] 1 2 3 4 5
But when you print that object you don’t see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help.) 
the "comment" attribute is a special kind that does not print, unlike the others

2. What happens to a factor when you modify its levels?
f1 <- factor(letters)
levels(f1) <- rev(levels(f1))
It reversed both the levels and the factor

3. What does this code do? How do f2 and f3 differ from f1?
f2 reverses the factor, but not the levels
f3 reverses the levels, but not the factor

  Matrices and Arrays
1. What does dim() return when applied to a vector?
NULL

2. If is.matrix(x) is TRUE, what will is.array(x) return?
TRUE, an array is a matrix

3. How would you describe the following three objects? What makes them different to 1:5?
x1 <- array(1:5, c(1, 1, 5)) a 1x1x5 array
x2 <- array(1:5, c(1, 5, 1)) a 1x5x1 array
x3 <- array(1:5, c(5, 1, 1)) a 5x1x1 array
They differ from 1:5 in that they have dimensions (3)

  Data Frames:
1. What attributes does a data frame possess?
names, row.names, class

2. What does as.matrix() do when applied to a data frame with columns of different types?
Coerces everything into the same type. 

3. Can you have a data frame with 0 rows? What about 0 columns?
Yes, an empty data frame

Subsetting:
  Data Types:

1. Fix each of the following common data frame subsetting errors:
mtcars[mtcars$cyl = 4, ] -> mtcars[mtcars$cyl == 4, ]
mtcars[-1:4, ] -> mtcars[-(1:4), ]
mtcars[mtcars$cyl <= 5] -> tcars[mtcars$cyl <= 5, ]
mtcars[mtcars$cyl == 4 | 6, ] -> mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]

2. Why does x <- 1:5; x[NA] yield five missing values? (Hint: why is it different from x[NA_real_]?)
X takes NA and expands it to a length of 5

3. What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
upper.tri() returns a logical matrix the size of the upper triangle of the matrix. Using it to subset the matrix, is in x[upper.tri(x)], returns a vector of the subsetted part of the matrix

4. Why does mtcars[1:20] return an error? How does it differ from the similar mtcars[1:20, ]?
mtcars[1:20] tells R to look at 20 columns, but there are not that many columns in mtcars. mtcars[1:20, ] tells R to look at rows 1:20 and all the columns

5. Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).

my_diag <- function(x){
  n <- nrow(x)
  return(x[matrix(seq_len(n), nrow = n, ncol = 2)])
}

6. What does df[is.na(df)] <- 0 do? How does it work?
subsets a dataframe to the values that are NA and turns NAs into 0s.

  Subsetting Operators

1. Given a linear model, e.g., mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Extract the R squared from the model summary (summary(mod))
mod$df.residual
summary(mod)$r.squared

  Applications

1.How would you randomly permute the columns of a data frame? (This is an important technique in random forests.) Can you simultaneously permute the rows and columns in one step?
df[,sample(ncol(df))]
df[sample(nrow(df)),sample(ncol(df))]

2. How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e., with an initial row, a final row, and every row in between)?
m <- some number 
testdf[sample(nrow(testdf), size = m), ]
testdf[sample(nrow(testdf) - m + 1, size = 1) + (0:(m - 1)), ]

3. How could you put the columns in a data frame in alphabetical order?
testdf[order(names(testdf))]







